---
title: "Practice Question 1"
author: "E P Swens"
subtitle: "Causal Inference"
output:
  pdf_document: default
  html_notebook: default
---

### A. Sketch the Directed Acyclic Graph (DAG)

Sketch (on paper or in `R`) the following DAG, representing our beliefs that:

- `x1 causes x2`
- `x1 causes x3`
- `x2 causes x3`

### B. Identify the number of open path(s) 

How many open path(s) are there?

A) 0
B) 1 
C) 2
D) 3

```{r include=FALSE}
# C
#
# X1 -> X3
# X1 -> X2 -> X3
```


### C. Identify the number of backdoor path(s) 

How many backdoor path(s) are there?

A) 0
B) 1 
C) 2
D) 3

```{r include=FALSE}
# A
```


### D. What is the valid adjustment set?

What are the valid adjustment set(s)? (Multiple solutions are possible)

A) $\emptyset$
B) $\{X_2\}$
C) $\{X_1, X_2\}$
D) $\{X_2, X_3\}$

```{r include=FALSE}
# A
```


### E. Simulate the Data

Simulate the data ($n = 1000$ with `set.seed(1)`) from the structural equations:  

$$X_1 \sim \mathcal{N}(0,1)$$
$$X_2 \sim 2 X_1 + \mathcal{N}(0,1)$$
$$X_3 \sim X_1 - 0.5X_2+ \mathcal{N}(0,1)$$

where $\epsilon_1, \epsilon_2, \epsilon_3 \sim \mathcal{N}(0,1)$ (i.i.d.)


```{r include=FALSE}
set.seed(1)

n <- 1000

x1 <- rnorm(n)
x2 <- 2 * x1 + rnorm(n)
x3 <- x1 + -0.5 * x2 + rnorm(n)

data <- data.frame(x1 = x1, x2 = x2, x3 = x3)
```


### F. Test the Correlations

Test the correlations between the variables use $\alpha = 0.05$. Tip: You can use `cor.test` from (`stats`) to test marginal dependencies and `pcor.test` from (`ppcor`) to test conditional dependencies. 

```{r include=FALSE}
library(ppcor)

cor.test(data$x1, data$x2)$p.value < 0.05
cor.test(data$x1, data$x3)$p.value < 0.05
cor.test(data$x2, data$x3)$p.value < 0.05

pcor.test(data$x1, data$x2, data$x3)$p.value < 0.05
pcor.test(data$x1, data$x3, data$x2)$p.value < 0.05
pcor.test(data$x2, data$x3, data$x1)$p.value < 0.05
```


### G. Recover the DAG

Which correlation(s) are (in)significant from part C?
What structure can you infer from the results?

A) Mediator
B) Collider
C) Confounder

```{r include=FALSE}
# B) Collider
```


### H. Recovered DAG from PC Algoritm

```{r echo=FALSE, fig.cap="DAG Recovered by PC Algoritm", fig.height=6, fig.width=10, message=FALSE, warning=FALSE, dpi=300}
library(pcalg)

pc <- pc(suffStat = list(C = cor(data), n = nrow(data)), 
         indepTest = gaussCItest,
         p = ncol(data), 
         alpha = 0.05)

plot(pc, main = "")
```

Figure 1 (next page) shows that the `PC-algoritm` did not recover the DAG. Which model assumption is not met? 

A) Positivity
B) Exchangeability 
C) Faithfulness
D) Sufficiency
E) Strong ignorability

```{r include=FALSE}
# C) Faithfulness
```


### I. Why does the algoritm require this assumption?

```{r include=FALSE}
# In order to infer d-separation from the conditional independence methods.
# Otherwise colliders cannot be identified from the conditional independence 
# methods
```


### J. In general, we do not worry about violation of this assumption, why?

```{r include=FALSE}
# Competitive mediation effects which result in a total effect of 0 are rare
```

















