---
title: "Practice Question 1"
author: "E P Swens"
subtitle: "Causal Inference"
output:
  pdf_document: default
  html_notebook: default
---

```{r echo=FALSE, fig.cap="Directed Acyclic Graph (Exposure X1, Outcome X5)", fig.height=4, fig.width=10, message=FALSE, warning=FALSE, dpi=300}
library(ggplot2)
library(dagitty)
library(ggpubr)
library(ggdag)

# DAG 1
dag1 <- dagify(
   X1 ~ X2,
   X2 ~ X1,
   X3 ~ X2 + X4,
   X5 ~ X1 + X3 + X4,
   exposure = "X1",
   outcome = "X5",
   coords = list(x = c(X1 = 0, X2 = 0, X3 = 1, X4 = 2, X5 = 2),
                 y = c(X1 = 0, X2 = 2, X3 = 1, X4 = 2, X5 = 0)))

f1 <- ggdag_status(dag1) + 
   guides(fill = "none", color = "none") + 
   theme_dag()

# DAG 2
dag2 <- dagify(
   X1 ~ X2,
   X3 ~ X2 + X4,
   X5 ~ X1 + X3 + X4,
   exposure = "X1",
   outcome = "X5",
   coords = list(x = c(X1 = 0, X2 = 0, X3 = 1, X4 = 2, X5 = 2),
                 y = c(X1 = 0, X2 = 2, X3 = 1, X4 = 2, X5 = 0)))

f2 <- ggdag_status(dag2) + 
   guides(fill = "none", color = "none") + 
   theme_dag()

# DAG 3
dag3 <- dagify(
   X1 ~ X5,
   X2 ~ X1,
   X3 ~ X2 + X4,
   X5 ~ X3 + X4,
   exposure = "X1",
   outcome = "X5",
   coords = list(x = c(X1 = 0, X2 = 0, X3 = 1, X4 = 2, X5 = 2),
                 y = c(X1 = 0, X2 = 2, X3 = 1, X4 = 2, X5 = 0)))

f3 <- ggdag_status(dag3) + 
   guides(fill = "none", color = "none") + 
   theme_dag()

# combined figure
ggarrange(f1, f2, f3, labels = c("A", "B", "C"), ncol = 3, nrow = 1)
```

### A. Directed Acyclic Graph

We are interested in the causal relationship between `x1` (exposure) and `x5` (outcome).

In figure 1, which diagram represents a directed acyclic graph?

A) A
B) B
C) C

```{r include=FALSE}
# B
```


### B. Continued 

Why are the other options not DAG?

```{r include=FALSE}
# A Bi-directional 
# C Cyclic
```


### C. Causal Paths

What is the number of path(s) in the DAG of figure 1?

A) 0
B) 1
C) 2
D) 3
E) 4

```{r include=FALSE}
# D

length(paths(dag2)$paths)
```

### D. Open Paths

How many of those path(s) are open?

A) 0
B) 1
C) 2
D) 3
E) 4

```{r include=FALSE}
# C

sum(paths(dag2)$open)
```


### E. Valid Adjustment Sets

Which set(s) adhere to the back-door criterion? (Multiple options are possible)

A) $\emptyset$
B) $\{X_2\}$
C) $\{X_3\}$
D) $\{X_4\}$
E) $\{X_2,X_3\}$
F) $\{X_2,X_4\}$
G) $\{X_3,X_4\}$
H) $\{X_2,X_3,X_4\}$

```{r include=FALSE}
# B, E, F, G, H

adjustmentSets(dag2, type = "all")
```
